### 2020-03-24 Paper_shortened_pdf.Rmd
Editor comments 1 and 2

The amount of time parents spend with children is associated with the quality of parenting behaviors [@sabattiniRelationMothersFathers2004]. Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a]. Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014]. Consequently, we hypothesize that greater joint involvement will positively attenuate both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively attenuate parental anger, $H_{4a-d}$.

By the same reasoning, we might hypothesize that greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004]. However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment. Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. One reason might be that parents are less able to monitor each other's child-rearing behavior when they are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Additionally, differential involvement might negatively impact the marital relationship, and thereby undermine both parents' involvement quality. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].

@@

We calculated mean scores for all study variables with multiple indicators. Scale descriptive statistics, number of items, attrition (indicated by the number of complete cases), and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2. The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents).

@@

## Correlations between qualitative parenting and emotional adjustment

We hypothesized that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment. At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)). Random intercepts reflect the average level of each variable across all waves of the study. Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment. In support of hypotheses $H_{1,2b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment. Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment. This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment. Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.

@@

We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)). In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between emotional adjustment and fathers' and mothers' warmth. In partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4). This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.

## Within-family predictive effects

At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment. The intercepts of these random effects reflect the average standardized regression coefficient across families. The variances in Table \@ref(tab:tabcrosspath) indicate whether this coefficient differs in strength (and sign) between families. We report unstandardized variances, because the standardized variances are, by definition, set to 1. With regard to within-families parenting effects, we found no evidence for hypotheses $H_{1b-d}$, as none of the intercepts differed significantly from zero. In line with hypothesis $H_{1a}$, however, on average, a maternal warmth positively predicted children's emotional adjustment in the next wave. The standardized regression coefficient of this effect, and others, is interpreted as follows: A $.99$ standard deviation increase in children's emotional adjustment in the next wave (see Table \@ref(tab:tabcrosspath)). The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.

@@

tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
tabparenteff$Parameter[c(1,4)] <- c("Random intercepts of parenting effects of:", "Variance of parenting effects of:")

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
tab <- rbind(tabparenteff, tabchildeff)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL

@@

tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
names(tab)[1] <- "Param."
reg_table(tab, caption = "Standardized between-family level intercepts and unstandardized variances of within-family predictive effects", note = "Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate. Matching letters indicate that parameters differed significantly; see Supplementary Table S4 for all difference tests.")
```

With regard to child effects, in line with hypothesis $H_{2a}$, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)). Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave. Contrary to hypothesis $H_{2b}$, the intercept of the effect of emotional adjustment on maternal consistency was non-significant. As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive. Counter to hypothesis $H_{2d}$, the intercept for maternal anger was positive, not negative. Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more anger than usual in the next wave. In support of hypothesis $H_{2a}$, the intercept for the child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts ($H_{2b-d}$) for child effects on paternal parenting were significant. The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.

@@

## Effects of quantitative involvement on parenting and adjustment

We examined standardized regression coefficients of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning. Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint involvement on parental anger, $H_{4d}$. With regard to relative involvement, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively. For mothers, the only significant association we found was that mothers displayed more warmth when they were relatively more involved than fathers. We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers. Counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment. Thus, children in families with highly involved parents generally displayed poorer emotional adjustment. Hypothesis $H_{6b}$ was supported: Children's emotional adjustment was lower in families where mothers were relatively more involved than fathers.

@@

To examine interactions between involvement quality and quantity, we examined standardized regression coefficients of between-family differences in quantitative involvement on within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth. Specifically, a one-standard deviation increase in joint involvement predicted a decrease in the within-families predictive effect of maternal warmth on emotional adjustment of $-.14$ of a standard deviation of the effect (see Table \@ref(tab:tabcrosspath)). This same effect was positively attenuated by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint quantitative involvement was higher, and also when mothers' relative involvement was greater compared to fathers. We found no effects for consistency. For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when joint involvement was higher. Emotional adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

@@

Evidence that quantitative moderates child effects was somewhat stronger. Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth. These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth. In families where mothers' involvement was high - either in terms of joint, or relative involvement - this positive effect was attenuated somewhat. Secondly, joint quantitative involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning. This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved. When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.

>>>>

The amount of time parents spend with children is associated with the quality of parenting behaviors [@sabattiniRelationMothersFathers2004]. Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a]. Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014]. Consequently, we hypothesize that greater joint involvement will positively predict both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively predict parental anger, $H_{4a-d}$.

By the same reasoning, we might hypothesize that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004]. However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment. Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. One reason might be that parents are less able to monitor each other's child-rearing behavior when they are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Additionally, differential involvement might negatively impact the marital relationship, and thereby undermine both parents' involvement quality. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].

@@

We calculated mean scores for all study variables with multiple indicators. Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2. The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents).

@@

## Is qualitative parenting correlated with emotional adjustment?

We hypothesized that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment. At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)). These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment. Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment. In support of hypotheses $H_{1,2b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment. Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment. Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment. This means that, in families where parents displayed more consistent parenting, children had better emotional adjustment - and in families where parents engaged in more inductive reasoning or anger, children had poorer emotional adjustment.

@@

We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)). These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in children's emotional adjustment. In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment. This means that if, in one particular wave, parents displayed more warmth than usual, their children tended to display better emotional adjustment than usual. Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4). This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.

## Predictive effects between within-family changes in parenting and emotional adjustment

In our analyses, each family has family-specific regression coefficients for the cross-lagged effects between wave-to-wave fluctuations in qualitative parenting, and emotional adjustment. Our hypotheses $H_{1a-d}$ also apply to these within-family predictive effects. To test these hypotheses, we examine the average standardized regression coefficients across families (see Table \@ref(tab:tabcrosspath)). In technical terms, these are random intercepts of the within-family predictive effects. Their variances indicate whether the family-specific regression coefficient differs in strength (and sign) across families. With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave. The standardized regression coefficient of this effect is interpreted as follows: On average across families, a one standard deviation increase in maternal warmth in one wave predicts a $.99$ standard deviation increase in children's emotional adjustment in the next wave. We found no evidence for hypotheses $H_{1b-d}$ in terms of parenting effects. The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.

@@

tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
tabparenteff$Parameter[c(1,4)] <- c("Average parenting effect of:", "Variance of parenting effect of:")

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Average child effect on:", "Variance of child effect on:")
tab <- rbind(tabparenteff, tabchildeff)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL

@@

tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
names(tab)[1] <- "Param."
reg_table(tab, caption = "Standardized between-family level intercepts (average effect) and unstandardized variances of within-family predictive effects.", note = "Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate. Matching letters indicate that parameters differed significantly; see Supplementary Table S4 for all difference tests.")
```

With regard to child effects, in line with hypothesis $H_{2a}$, the average cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)). Thus, on average, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave. Contrary to hypothesis $H_{2b}$, the average effect of emotional adjustment on maternal consistency was non-significant. As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive. Counter to hypothesis $H_{2d}$, the intercept for maternal anger was positive, not negative. Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more anger than usual in the next wave. In support of hypothesis $H_{2a}$, the average child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts ($H_{2b-d}$) for child effects on paternal parenting were significant. The variances of all child effects were significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.

@@

## Does quantitative involvement predict the average level of parenting and adjustment?

To investigate our hypotheses $H_{4-6}$, we examined the predictive effects of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to effects of joint quantitative involvement on parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, greater joint involvement predicted increased warmth, consistency, and inductive reasoning. Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint involvement on parental anger, $H_{4d}$. With regard to effects of relative quantitative involvement on parenting, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent. For mothers, the only significant association we found was that mothers displayed more warmth when they were relatively more involved than fathers. We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers. With regard to effects of quantitative involvement on emotional adjustment, counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment. Thus, children of highly involved parents generally displayed poorer emotional adjustment. Hypothesis $H_{6b}$ was supported: Children's emotional adjustment was lower in families where mothers were relatively more involved than fathers.

@@

To examine interactions between involvement quality and quantity, we investigated whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth. Specifically, a one-standard deviation increase in joint involvement predicted a $-.14$ standard deviation decrease in the within-families predictive effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)). This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that parents' greater joint quantitative involvement, and mothers' greater relative involvement, both negatively predicted the within-family effect of adjustment on maternal warmth. We found no effects for consistency. For inductive reasoning, we found that greater joint involvement negatively predicted the within-family effect of adjustment on maternal inductive reasoning. Emotional adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

@@

Evidence that quantitative moderates child effects was somewhat stronger. Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth. These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth. This positive effect was somewhat smaller in families where mothers' involvement was high - either in terms of joint, or relative involvement. Secondly, joint quantitative involvement positively moderated the link from emotional adjustment to maternal inductive reasoning. This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved. When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.
